### 2016-12-20 coverletter.tex

Dear Editor,

en el contexto de nuestra labor en el bci lab en el departamento bla bla b

This submitted article encompass the work that we have been doing and in the context of the setup of a Brain Computer Interfaces Laboratory at the Computer Engineering Department of the ITBA University in Buenos Aires Argentina.

Son areas de investigacion en las cuales nosotros podemos llegar a contribuir como ingenieros e investigadores a los efectos de brindar nuevos metodos y herramientas que favorezcan la comunicación o el tratamiento de individuos que han sido dañados o afectados por enfermedades neuro degenerativas

EEG Processing and BCI are research areas with amazing prospects for the general population but also to very specific population groups that require all the help that we can provide, as engineers and researchers.

We believe that this proposed method based on the histograms of gradients of signal plots, and most importantly, its successful application to identify ERP markers within EEG signals from ALS patients is relevant for this special issue and that it will contribute with a novel tool to analyse EEG signals.

Yours truly and deeply greatfull for your help
Rodrigo Ramele
PhD Candidate
ITBA University 
Argentina